Augmentation of a Historical Harpsichord Keyboard Replica for Haptic-Enabled Interaction in Museum Exhibitions

### 2025-02-03 src/article.tex
fix #39 #40

The keyboard presented here inherits some aspects from the Tromba Moderna project. The project aims to offer a tool to enhance the fruition of the \anon{Tagliavini} collection while retaining a form of continuity with historical instrument-building traditions. However, the electronics hidden inside the interface are not intended to augment or disrupt the tactile feedback; rather, they serve as the silent and invisible link between the mechanical and the digital realms. 
The optical sensing technique for the keyboard is adapted from a similar project on the piano by McPherson \cite{McPherson2013}. 
Thus, whether the interface proposed here is a \emph{new} musical interface is subject to discussion, particularly regarding the keyboard's tactile response designed to adhere to longstanding harpsichord building traditions. This work complements and follows up on previous reflections within the NIME community, emphasising the \emph{O} in NIME \cite{Masu_NIME_2023}. The intended use of the keyboard through meaningful interaction with a museum exhibit is where the novelty of this work lies, rather than solely in its technological development. Relying on previous NIMEs, this project extends the longevity of the results beyond the scope envisaged in the original works, enhancing their sustainability, as discussed by Masu \emph{et al.} \cite{Masu_NIME_2023}.

@@

\begin{figure*}
\centering
\includegraphics[width=0.8\linewidth]{src/images/details.jpg}
\caption{Details of the replica, including the key slots and sides, keys, and jacks with seagull plectra.}\label{fig:details}
\end{figure*}

@@

The design of the augmented replica keyboard for the \anon{Tagliavini} Collection was guided by three principal constraints stipulated by the museum:
\begin{itemize}
\item \emph{No visible electronics}. To preserve the visual integrity of the exhibit, all electronic components had to remain concealed except for a pair of headphones and a small display for audio parameter adjustments.
\item \emph{Robustness and reliability}. The system needed to accommodate frequent use by museum visitors and allow for straightforward maintenance by staff without requiring specialised technical expertise.
\item \emph{Faithfulness}. The keyboard mechanism ensured fidelity to its authentic operation, keeping artefacts to a minimum. 
\end{itemize}

These constraints demanded a non-invasive design that was respectful of the historical authenticity of the instruments. The requirement for faithfulness to the original mechanism immediately ruled out a mechanical design reliant on electromechanical actuators, as in previous works on piano haptics \cite{Timmermans2020,Gillespie1996}. 
While actuators can generate considerable force, it is difficult to achieve a jack's free motion and resistance without extensive mechanical intervention such as that carried out by Gillespie \cite{Gillespie1996}. The instrument itself can already achieve the tactile realism of the harpsichord. 
It should first be determined if mechanical recreation is sufficient as part of a digital musical instrument before pursuing elaborate electromechanical emulation. 
% (A REFERENCE WOULD BE GOOD HERE, ANY IDEAS?)
% The final design, visible in Figures \ref{fig:teaser} and \ref{fig:details}, is a 49-key two-register harpsichord keyboard replica.
% The keyboard was consciously chosen to mimic the typical layout of 16th-century Italian harpsichords. This design decision was made to use humans' propensity to being influenced by visual components when musical making judgements \cite{Tsay2013} to our advantage. It is especially true concerning musical instruments, as the studies by Fritz \emph{et al}\cite{Fritz2012, Fritz2014, Fritz2017} have shown. Rather than avoiding this effect, we have chosen to utilise it instead. This `musical instrument McGurk effect' is partly the reason behind hiding the electronics out of sight. 
% The aesthetics of the interface make it more likely that the interface be accepted on the same terms as an `authentic' musical instrument. With all the work into modelling and recreating piano action \cite{Cadoz1990, Gillespie1996, Timmermans2020} it is possible the visual component may provide the necessary `push' of acceptance similar to that which influences our judgment on performance \cite{Tsay2013}.

The final design, shown in Figures \ref{fig:teaser} and \ref{fig:details}, is a 49-key, two-register harpsichord keyboard replica. The keyboard layout was deliberately modelled after typical 16th-century Italian harpsichords, leveraging the human tendency to be influenced by visual elements when making musical judgments \cite{Tsay2013}. This effect is particularly relevant in musical instruments, as demonstrated by studies conducted by Fritz \emph{et al.} \cite{Fritz2012, Fritz2014, Fritz2017}. Instead of attempting to mitigate this influence, the design embraces it. This phenomenon, akin to a `musical instrument McGurk effect’, is one of the reasons why the electronics have been concealed from view. \textcolor{red}{Any good references on the McGurk effect?} The aesthetics of the interface enhance its likelihood of being perceived as an `authentic' musical instrument. Given the extensive research into modelling and recreating piano action \cite{Cadoz1990, Gillespie1996, Timmermans2020}, the visual component may provide the final persuasive element necessary for acceptance, similar to the way visual perception affects judgments of musical performance \cite{Tsay2013}.

@@

Optical sensors are positioned in front of each jack and their output voltage depends on the amount of light reflected from a sensor surface. In this case, the sensor surface is a grey-scale gradient printed on a vinyl sticker (Figure \ref{fig:jack-tags}) and applied to the side of each jack. Though the current configuration simply sends MIDI note-on and -off messages when passing a threshold, the displacement of all jacks is available continuously.
The staggering of the pluck point of the jacks between registers means that there are two distinct tactile anchors along the key dip. A combination of string tension, quill voicing, and jack staggering contribute to the overall `weight' of the action, with each key differing from the low to the high octaves \cite{Veroli2012}. Such tactile expectations provide fertile opportunities for extending the expressive capabilities of the interface. The current sensor system enables the data recording of traditional use from which it can be identified where those opportunities for new expression lie \cite{McPherson2013-2}.

@@

@@

A further objective, set forth by the authors, was to ensure accessibility and reproducibility by committing to an open source approach for all outputs of the project. The commitment to open sourcing encompassed all aspects of the system, including hardware schematics, firmware, and calibration data. Cost-effectiveness was also a central consideration. 
% The system was initially developed for a 3-key prototype, shown in Figure \ref{fig:3key}, and successfully scaled to 49 keys without significant increases in cost or complexity. 
Specifically, the system was designed to be easily assembled using resources typically available in a university-managed maker space.

Components, such as QRE1113 optical sensors and CD4051BE multiplexers, are widely available from commercial resellers, while the modular PCB design ensures easy replication and maintenance.

The Arduino Nano 33 BLE was chosen as the core microcontroller for its compatibility with open source tools and ability to support both USB and BLE MIDI. Ferroelectric RAM (FRAM) was employed for data storage since it offers an economical yet robust solution for preserving calibration settings across power cycles. Calibration workflows were optimised using the Arduino IDE’s serial plotter and open source MIDI Monitor software \footnote{\url{https://github.com/krevis/MIDIApps}}, reducing reliance on proprietary tools and simplifying the process for users.

@@

\subsection{Materials and Construction}

The keyboard was designed to replicate the tactile and aesthetic sensations of playing an antique Italian harpsichord. Traditional materials were used, including walnut for the wrest plank, chestnut for the key levers, boxwood and ebony for key covers, and cypress for the case and soundboard. The 98 jacks were made from beech, fitted with brass springs and natural seagull feather plectra. The design was inspired by Venetian harpsichords, particularly the \anon{1547 Alessandro Trasuntino} harpsichord at the \anon{San Colombano Museum}. An exception was made for the short octave --- the assigning of common keys in the first octave instead of a chromatic scale --- typically found in Italian harpsichords, which was replaced by a standard octave layout to more easily accommodate the interaction with commercial sample libraries.

The rectangular poplar frame also deviates from the traditional logarithmic \ref{fig:log-harp-comp} form to allow for the installation of the electronic sensors, as visible in Figure \ref{fig:49-key-bottom}, without compromising the visual or tactile authenticity. Two string choirs, crafted from yellow brass wire and anchored with wrought iron pins, were tensioned to replicate authentic plucking resistance. Felt strips were added to dampen vibrations. The result is an interface that combines the mechanical action of keys with synthetic sound generation, preserving a real harpsichord's tactile qualities.

@@

@@

\begin{figure}
    \centering    
    \includegraphics[width=\linewidth]{src/images/3-key-side.png}
    \caption{3-Key Model Harpsichord Mechanism acquired from \anon{San Colombano, Bologna}, 2023}
    \label{fig:3key}
\end{figure}

@@

\begin{itemize}
    \item \emph{Non-invasiveness.} No remarkable modifications were allowed on the harpsichord mechanics, particularly on all the visible parts.
    \item \emph{Low Latency.} The sampling period for reading and processing data from all sensors should remain under 10 ms to allow for latency introduced in other parts of the synthesis process. Empirical criteria found in previous studies \cite{Jack2016} was used as a guide.
    \item \emph{Reliability:} Sensor data should be dependable and consistent, with interference from anything external to the jack movement being absent or negligible.
    \item \emph{Scalability.} The design needed to scale in both cost and time in order to be adaptable from a 3-key prototype (Figure \ref{fig:3key}) up to the 49 keys of the final design.
    % \item \emph{Expandability.} The system should accommodate future functionality, such as additional MIDI parameters or data visualisation.
\end{itemize}

The hardware was designed to make assembly possible in standard maker spaces.

@@

\begin{figure}[!h] 
  \centering
  \includegraphics[width=0.7\linewidth,trim={0 1.5cm 0 1.5cm},clip]{src/images/simple-schematic-bw-.jpg} 
  \caption{Optical sensor in a simple voltage divider circuit. \texttt{V\_OUT} is routed to one of 8 channels on the CD4051BE multiplexer}
  \Description{} 
  \label{fig:simple-schematic}
\end{figure}

@@

The optical sensors are wired in a voltage divider configuration \ref{fig:simple-schematic} and the circuit outputs a voltage based on the infrared light reflected from nearby surfaces. 
The gradient stickers affixed to each jack provided a surface with varying reflectivity for the sensors, which were used to track jack displacement throughout the key dip.

@@

\begin{itemize}
    \item 1 Arduino Nano 33 BLE
    \item 1 Fujitsu MB85RS64 SPI FRAM chip
    \item 1 EC11 combined rotary encoder and tactile switch
    % \item 1 DFROBOT DFR0785 Tactile Switch
\end{itemize}

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{src/images/block-diagram-2.pdf}
    \caption{Block diagram of PCB connections. Sensor and LED are routed through each sensor PCB}
    \label{fig:system-block-diagram}
\end{figure*}

The Arduino Nano small form factor made fitting the board inside the harpsichord easier and also allowed for testing of multiple chipsets. The 33 BLE variation of the Nano had the additional benefit in its ability to be programmed as a BLE MIDI device.  
%Initial iterations of the controller board used 2.54 mm pitch headers for cable connections, but issues with cable disconnections led to the adoption of JST-PH connectors in later designs. Custom cable looms were constructed using adapted crimping tools, ensuring robust connections without compromising accessibility.

Non-volatile memory, implemented using a Fujitsu MB85RS64 Ferroelectric RAM (FRAM) chip, provided a reliable means of storing calibration data. This ensured that sensor thresholds and other settings could be preserved across power cycles, enhancing the system's usability in museums.

%\subsection{Calibration and Power Management}\label{calibration}

@@

% \begin{figure}
%     \centering
%     \includegraphics[width=\linewidth]{src/images/serial_monitor.png}
%     \caption{Arduino IDE serial plotter used for calibration of sensor thresholds. \texttt{M} and \texttt{m} represent the maximum an minimum value of digitised data. \texttt{K} and \texttt{T} are the values of the sensor and threshold being calibrated, followed by a two-digit number denoting the sensor index (in this case, sensor 32).}
%     \label{fig:serial_monitor}
% \end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth,trim={0 5cm 0 0},clip]{src/images/mac-mini.jpg}
    \caption{Components and cabling hidden in the bottom section of the instrument case.}
    \label{fig:mac-mini}
\end{figure}

% \subsection{Final Assembly and Integration}

% The final 49-key harpsichord replica incorporated the sensor and controller boards into the rectangular frame. Ribbon cables connected the PCBs, while felt strips were interwoven between the strings to suppress vibrations and maintain the tactile response of the keys. The assembly process at the \anon{NEMUS} lab in \anon{Bologna} involved meticulous adjustments to key clearances and wiring layouts. Modifications included milling sections of the keys to ensure adequate clearance for the PCBs and varnishing the jacks to improve sticker adhesion.

For the initial version of the exhibition, a harpsichord sample library was used and controlled via Native Instrument's Kontakt\footnote{\href{https://www.native-instruments.com/en/products/komplete/samplers/kontakt-8/?srsltid=AfmBOorKUf43SoIxGBS2-GnXmKHHkgcfcfWRskpweDhLSG3FiF0qrf2w}{Kontakt webpage} (accessed 31 Jan 2025)}. The software was installed on a Mac Mini hosted inside the instrument's case. Holes were drilled at the back of the instrument's frame, away from sight, to allow the passage of cables, such as from power supplies, USB connectors and headphone jack (Figure \ref{fig:mac-mini}). 
% These components are accessible via locked drawer which is visible in \ref{fig:oratory}. 
An iPad is used as a monitor through which visitors can adjust playback parameters such as tuning, voicing and which stops are engaged. The next iteration will use a library of samples from instruments in the collection, which the museum is currently carrying out.
%A main 12-inch display sits inside a custom-made drawer locked during usage.

\begin{figure}[!b]
    \centering
    \includegraphics[width=\linewidth,trim={0 1.5cm 0 1.5cm},clip]{src/images/sensor-board-w-baffles.jpeg}
    \\
    \includegraphics[width=0.8\linewidth,trim={0 2cm 0 2.5cm},clip]{src/images/baffles.png}
    \caption{Baffles designed to prevent cross-talk between adjacent sensors.}
    \Description{}
    \label{fig:baffles}
\end{figure}

@@

% In an effort to make the exhibition more accessible and authentic, the \anon{San Colombano Museum} commissioned the construction of a mechanical keyboard to trigger a MIDI-controlled sample library, designed for use by museum visitors. 
The exhibition is set to open approximately a month after this work's submission, and only preliminary feedback has been collected from a pool of 20 visitors -- who were given training on the system -- and expert feedback from \anon{Catalina Vicens} and \anon{Roberto Livi}.
The collection of visitor feedback was carried out as short informal interviews and primarily as a means of identifying any technical problems.
The keyboard shows promise in enhancing the museum experience as comments about a ``sense of disconnect'' found in the Benton Fletcher Collection \cite{McAlpine2014} were absent in this initial feedback. \textcolor{red}{This goes in the direction pointed by Andrew, but I wonder if it could be worded differently}

>>>>

The keyboard presented here inherits some aspects from the Tromba Moderna project. The project aims to offer a tool to enhance the fruition of the \anon{Tagliavini} collection while retaining a form of continuity with historical instrument-building traditions. However, the electronics hidden inside the interface are not intended to augment or disrupt the tactile feedback; rather, they serve as the silent and invisible link between the mechanical and the digital realms. 
The optical sensing technique for the keyboard is adapted from a similar project on the piano by McPherson \cite{McPherson2013}. 
Thus, whether the interface proposed here is a \emph{new} musical interface is subject to discussion, particularly regarding the keyboard's tactile response designed to adhere to longstanding harpsichord building traditions. This work complements and follows up on previous reflections within the NIME community, emphasising the \emph{O} in NIME -- the methodologies and reasoning for exploring `old' instrument designs at NIME \cite{Masu_NIME_2023}. The intended use of the keyboard through meaningful interaction with a museum exhibit is where the novelty of this work lies, rather than solely in its technological development. Relying on previous NIMEs, this project extends the longevity of the results beyond the scope envisaged in the original works, enhancing their sustainability, as discussed by Masu \emph{et al.} \cite{Masu_NIME_2023}.

@@

\begin{figure*}
\centering
\includegraphics[width=0.8\linewidth]{src/images/details.jpg}
\caption{Details of the replica, including the key slots and sides, keys, and jacks with seagull plectra.}
\label{fig:details}
\end{figure*}

@@

The design of the augmented replica keyboard for the \anon{Tagliavini} Collection was guided by three principal constraints stipulated by the museum:

\begin{itemize}
    \item \emph{No visible electronics}. To preserve the visual integrity of the exhibit, all electronic components had to remain concealed except for a pair of headphones and a small display for audio parameter adjustments.
    \item \emph{Robustness and reliability}. The system needed to accommodate frequent use by museum visitors and allow for straightforward maintenance by staff without requiring specialised technical expertise.
    \item \emph{Faithfulness}. The keyboard mechanism ensured fidelity to its authentic operation. The electronics system would not seek to `fix' or `improved' on limitations of the original 16th-century design. 
\end{itemize}

These constraints demanded an electronics system that was non-invasive in design and that was respectful of the historical aesthetics of the instruments. The requirement for faithfulness to the original mechanism immediately ruled out a mechanical design reliant on electromechanical actuators, as in previous works on piano haptics \cite{Timmermans2020,Gillespie1996}. 
While actuators can generate considerable force, it is difficult to achieve a jack's free motion and resistance without extensive mechanical intervention such as that carried out by Gillespie \cite{Gillespie1996}. A recreation of the mechanism can already achieve the tactile realism of the harpsichord.
It should first be determined if mechanical recreation is sufficient as part of a digital musical instrument before pursuing elaborate electromechanical emulation.

The final design, shown in Figures \ref{fig:teaser} and \ref{fig:details}, is a 49-key, two-register harpsichord keyboard replica. The keyboard layout was deliberately modelled after typical 16th-century Italian harpsichords, leveraging the human tendency to be influenced by visual elements when making musical judgments \cite{Tsay2013}. This effect is particularly relevant in musical instruments, as demonstrated by studies conducted by Fritz \emph{et al.} \cite{Fritz2012, Fritz2014, Fritz2017}. Instead of attempting to mitigate this influence, the design embraces it. This phenomenon, akin to a `musical instrument McGurk effect’, is one of the reasons why the electronics have been concealed from view. The aesthetics of the interface enhance its likelihood of being perceived as an `authentic' musical instrument. Given the extensive research into modelling and recreating piano action \cite{Cadoz1990, Gillespie1996, Timmermans2020}, the visual component may provide the final persuasive element necessary for acceptance, similar to the way visual perception affects judgments of musical performance \cite{Tsay2013}.

@@

Optical sensors are positioned in front of each jack and their output voltage depends on the amount of light reflected from a sensor surface. In this case, the sensor surface is a grey-scale gradient printed on a vinyl sticker (Figure \ref{fig:jack-tags}) and applied to the side of each jack. Though the current configuration simply sends MIDI note-on and -off messages when passing a threshold, the displacement of all jacks is available continuously.
The jacks for a single key typically have their pluck points offset from each other, a practice known as `staggering'.
A combination of string tension, quill voicing, and jack staggering contribute to the overall `weight' of the action, with each key differing from the low to the high octaves \cite{Veroli2012}. The staggering of the pluck point of the jacks between registers means that there are two distinct tactile anchors along the key dip. The current sensor system enables the data recording of traditional use from which it can be identified where those opportunities for new expression lie \cite{McPherson2013-2}.

@@

\subsection{Materials and Construction}
The keyboard was designed to replicate the tactile and aesthetic sensations of playing an antique Italian harpsichord. Traditional materials were used, including walnut for the wrest plank, chestnut for the key levers, boxwood and ebony for key covers, and cypress for the case and soundboard. The 98 jacks were made from beech, fitted with brass springs and natural seagull feather plectra. The design was inspired by Venetian harpsichords, particularly the \anon{1547 Alessandro Trasuntino} harpsichord at the \anon{San Colombano Museum}. An exception was made for the short octave --- the assigning of common keys in the first octave instead of a chromatic scale --- typically found in Italian harpsichords, which was replaced by a standard octave layout to more easily accommodate the interaction with commercial sample libraries.

The rectangular poplar frame also deviates from the traditional logarithmic (Figure \ref{fig:log-harp-comp}) form to allow for the installation of the electronic sensors, as visible in Figure \ref{fig:49-key-bottom}, without compromising the visual or tactile authenticity. Two string choirs, crafted from yellow brass wire and anchored with wrought iron pins, were tensioned to replicate authentic plucking resistance. Felt strips were added to dampen vibrations. The result is an interface that combines the mechanical action of keys with synthetic sound generation, preserving a real harpsichord's tactile qualities.

@@

A further objective, set forth by the authors, was to ensure reproducibility by committing to an open source approach for all outputs of the project. The commitment to open sourcing encompassed all aspects of the system, including hardware schematics, firmware, and calibration data. Cost-effectiveness was also a central consideration. 
% The system was initially developed for a 3-key prototype, shown in Figure \ref{fig:3key}, and successfully scaled to 49 keys without significant increases in cost or complexity. 
Specifically, the system was designed to be easily assembled using resources typically available in a university-managed maker space. Components, such as QRE1113 optical sensors and CD4051BE multiplexers, are widely available from commercial resellers, while the modular PCB design ensures easy replication and maintenance. The Arduino Nano was chosen as the core microcontroller format for its compatibility with open source tools. Calibration workflows were optimised using the Arduino IDE’s serial plotter and open source MIDI Monitor software \footnote{\url{https://github.com/krevis/MIDIApps}}, reducing reliance on proprietary tools and simplifying the process for users.

@@

@@

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{src/images/block-diagram-2.pdf}
    \caption{Block diagram of PCB connections. A separate sensor signal is routed to the Arduino. LED and multiplexer (MUX) controls signal are daisy-chained through each sensor PCB.}
    \label{fig:system-block-diagram}
\end{figure*}

@@

\begin{figure}
    \centering    
    \includegraphics[width=\linewidth]{src/images/3-key-side.png}
    \caption{3-Key Model Harpsichord Mechanism acquired from \anon{San Colombano, Bologna}, 2023.}
    \label{fig:3key}
\end{figure}

@@

\begin{itemize}
    \item \emph{Non-invasiveness:} No remarkable modifications were allowed on the harpsichord mechanics, particularly on all the visible parts.
    \item \emph{Low Latency:} The sampling period for reading and processing data from all sensors should remain under 10 ms to allow for latency introduced in other parts of the synthesis process. Empirical criteria found in previous studies \cite{Jack2016} was used as a guide.
    \item \emph{Reliability:} Sensor data should be dependable and consistent, with interference from anything external to the jack movement being absent or negligible.
    \item \emph{Scalability:} The design needed to scale in both cost and time in order to be adaptable from a 3-key prototype (Figure \ref{fig:3key}) up to the 49 keys of the final design.
    % \item \emph{Expandability:} The system should accommodate future functionality, such as additional MIDI parameters or data visualisation.
\end{itemize}

The hardware was designed to make assembly possible in standard university maker spaces.

@@

\begin{figure}[!h] 
  \centering
  \includegraphics[width=0.7\linewidth,trim={0 0.5cm 0 0.5cm},clip]{src/images/simple-schematic-bw-.jpg} 
  \caption{Optical sensor in a simple voltage divider circuit. \texttt{V\_OUT} is routed to one of 8 channels on the CD4051BE multiplexer.}
  \Description{} 
  \label{fig:simple-schematic}
\end{figure}

@@

The optical sensors are wired in a voltage divider configuration (Figure \ref{fig:simple-schematic}) and the circuit outputs a voltage based on the infrared light reflected from nearby surfaces. 
The gradient stickers affixed to each jack provided a surface with varying reflectivity for the sensors, which were used to track jack displacement throughout the key dip.

\begin{figure}[!t]
    \centering
    \includegraphics[width=\linewidth,trim={0 1.5cm 0 1.5cm},clip]{src/images/sensor-board-w-baffles.jpeg}
    \\
    \includegraphics[width=0.8\linewidth,trim={0 2cm 0 2.5cm},clip]{src/images/baffles.png}
    \caption{Baffles designed to prevent cross-talk between adjacent sensors.}
    \Description{}
    \label{fig:baffles}
\end{figure}

@@

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.8\linewidth,trim={0 5cm 0 0},clip]{src/images/mac-mini.jpg}
    \caption{Components and cabling hidden in the bottom section of the instrument case.}
    \label{fig:mac-mini}
\end{figure}

\begin{itemize}
    \item 1 Arduino Nano 33 BLE.
    \item 1 Fujitsu MB85RS64 SPI Ferroelectric RAM chip.
    \item 1 EC11 combined rotary encoder and tactile switch.
    % \item 1 DFROBOT DFR0785 Tactile Switch
\end{itemize}

The Arduino Nano's small form factor made fitting the board inside the harpsichord easier and also allowed for testing of multiple chipsets. The 33 BLE variation of the Nano had the additional benefit in its ability to be programmed as both native USB MIDI and BLE MIDI devices. The 33 BLE was also able to achieve a sampling period of 1.7 milliseconds for all sensors. A 4-sample moving average filter was implemented to reduce noise and brought the effective sampling period to 5.1 milliseconds.
% Samping period was measure by timing each iterations of the entire sensor read and processing cycle, polling the internal clock of the arduino a the start and end of each cycle and storing the difference in a buffer.  After 10k cylces the buffer of timing was sent over serial and a mean average of 1.7 milliseconds was found with a jitter of ±0.2 milliseconds.

Non-volatile memory, implemented using a Fujitsu MB85RS64 Ferroelectric RAM (FRAM) chip, provided a reliable means of storing and preserving calibration settings across power cycles. This ensured that sensor thresholds and other settings could be preserved across power cycles, enhancing the system's usability in museums.

@@

For the initial version of the exhibition, a harpsichord sample library was used and controlled via Native Instrument's Kontakt\footnote{\href{https://www.native-instruments.com/en/products/komplete/samplers/kontakt-8/?srsltid=AfmBOorKUf43SoIxGBS2-GnXmKHHkgcfcfWRskpweDhLSG3FiF0qrf2w}{Kontakt webpage} (accessed 31 Jan 2025)}. The software was installed on a Mac Mini hosted inside the instrument's case. Holes were drilled at the back of the instrument's frame, away from sight, to allow the passage of cables, such as from power supplies, USB connectors and headphone jack (Figure \ref{fig:mac-mini}). 
An iPad is used as a monitor through which visitors can adjust playback parameters such as tuning, voicing and which stops are engaged. The next iteration will use a library of samples from instruments in the collection, which the museum is currently carrying out.

@@

% In an effort to make the exhibition more accessible and authentic, the \anon{San Colombano Museum} commissioned the construction of a mechanical keyboard to trigger a MIDI-controlled sample library, designed for use by museum visitors. 
The exhibition is set to open approximately a month after this work's submission, and only preliminary feedback has been collected from a pool of 20 visitors -- who were given training on the system -- and expert feedback from \anon{Catalina Vicens} and \anon{Roberto Livi}.
The collection of visitor feedback was carried out as short informal interviews and primarily as a means of identifying any technical problems.
The keyboard shows promise in enhancing the museum experience as comments about a ``sense of disconnect'' found in the Benton Fletcher Collection \cite{McAlpine2014} were absent in this initial feedback.

%\textcolor{red}{This goes in the direction pointed by Andrew, but I wonder if it could be worded differently}
